Modeling the impacts of park access on health outcomes: a utility-based accessibility approach

### 2020-09-25 manuscript/manuscript.Rmd
Add sentence about putting amenities into utility in contribution
Also, minor typo correction

To our knowledge, there has been no previous attempt to examine
access to parks in a metropolitan area using a utility-based definition as the
primary indicator of access. Consequently, there has also been no previous
effort to examine the effect of park access on health outcomes using this
theoretically attractive technique. In this study, we first develop a utility-based
measure of park access in New York City, and then apply this measure to study the
effect of access to public parks on aggregate health outcomes.

@@

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
access and public health, previous literature has produced mixed
findings, perhaps owing to the wide range of accessibility definitions. 
Using New York City as a case study, we presented a
holistic and flexible measurement for park accessibility that compares the
continuous distance to all parks in the region, weighted against the size of the
park and its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive relationship
where the least park-accessible tracts have an expected physical activity
participation rate between `r 6 * .01` and `r 6 * .3` percentage points lower than the most
accessible tracts. We also found a suggestive though not significant relationship
between park accessibility and expected obesity rates *in addition to* the effect
of physical activity participation.

>>>>

To our knowledge, there has been no previous attempt to examine
access to parks in a metropolitan area using a utility-based definition as the
primary indicator of access. Consequently, there has also been no previous
effort to examine the effect of park access on health outcomes using this
theoretically attractive technique. In this study, we first develop a utility-based
measure of park access in New York City, and then apply this measure to study the
effect of access to public parks on aggregate health outcomes. In this
application, we demonstrate how additional park amenities including park size,
sport facilities, playgrounds, and trails can be incorporated in a more complete
and informative measure of access beyond simple proximity.

@@

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
access and public health, previous literature has produced mixed
findings, perhaps owing to the wide range of accessibility definitions. 
Using New York City as a case study, we presented a
holistic and flexible measurement for park accessibility that compares the
continuous distance to all parks in the region, weighted against the size of the
park and its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive relationship
where the least park-accessible tracts have an expected physical activity
participation rate between $0.3$ and $2.1$ percentage points lower than the most
accessible tracts. We also found a suggestive though not significant relationship
between park accessibility and expected obesity rates *in addition to* the effect
of physical activity participation.

### 2020-09-25 manuscript/manuscript.tex
Add sentence about putting amenities into utility in contribution
Also, minor typo correction

\begin{table*}
\caption{Comparison of SLX and SDEM Coefficients}
\label{tab:slx-sdem}
\begin{tabular}{l d{2.4} d{2.4} d{2.4} c d{2.4} d{2.4} d{2.4} }
\toprule
 & \multicolumn{3}{c}{SLX} & & \multicolumn{3}{c}{SDEM} \\
\cmidrule{2-4}\cmidrule{6-8}
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.} & 
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.}\\
\midrule
% latex table generated in R 4.0.0 by xtable 1.8-4 package
% Thu Sep 24 14:12:00 2020
 (Intercept) & -15.8791 & -26.8066 & -4.9515 &   & -24.7352 & -41.5391 & -7.9312 \\ 
  log(Density) & 0.0502 & -0.1280 & 0.2284 &   & 0.1760 & 0.0371 & 0.3150 \\ 
  log(Income) & 5.8667 & 5.3476 & 6.3859 &   & 6.0134 & 5.5901 & 6.4368 \\ 
  Fulltime & 0.1157 & 0.0944 & 0.1369 &   & 0.1271 & 0.1086 & 0.1455 \\ 
  College-educated & 0.0029 & -0.0253 & 0.0311 &   & 0.0077 & -0.0156 & 0.0310 \\ 
  Single Adults & -0.0435 & -0.0623 & -0.0247 &   & -0.0365 & -0.0528 & -0.0201 \\ 
  Youth (0-17) & -0.1528 & -0.1844 & -0.1211 &   & -0.1320 & -0.1585 & -0.1055 \\ 
  Young adults (18-34) & 0.0245 & -0.0001 & 0.0492 &   & 0.0292 & 0.0092 & 0.0493 \\ 
  Seniors (65+) & 0.0108 & -0.0212 & 0.0427 &   & 0.0386 & 0.0112 & 0.0659 \\ 
  Black population share & -0.0471 & -0.0589 & -0.0353 &   & -0.0499 & -0.0586 & -0.0412 \\ 
  Asian population share & -0.0684 & -0.0828 & -0.0540 &   & -0.0764 & -0.0870 & -0.0659 \\ 
  Hispanic population share & -0.1003 & -0.1126 & -0.0881 &   & -0.1023 & -0.1114 & -0.0932 \\ 
  Other Minorities & 0.0341 & -0.0661 & 0.1343 &   & 0.0047 & -0.0837 & 0.0931 \\ 
  $\gamma$: log(Density) & 1.2513 & 0.9367 & 1.5659 &   & 1.0873 & 0.7062 & 1.4684 \\ 
  $\gamma$: log(Income) & 2.3641 & 1.5266 & 3.2016 &   & 1.9780 & 0.9563 & 2.9997 \\ 
  $\gamma$: Fulltime & -0.1346 & -0.1749 & -0.0943 &   & -0.0411 & -0.0882 & 0.0059 \\ 
  $\gamma$: College-educated & -0.1201 & -0.1692 & -0.0711 &   & -0.0704 & -0.1296 & -0.0112 \\ 
  $\gamma$: Single Adults & -0.0520 & -0.0857 & -0.0183 &   & -0.0094 & -0.0514 & 0.0325 \\ 
  $\gamma$: Youth (0-17) & -0.1013 & -0.1594 & -0.0432 &   & -0.0128 & -0.0815 & 0.0559 \\ 
  $\gamma$: Young adults (18-34) & 0.1131 & 0.0706 & 0.1555 &   & 0.0838 & 0.0334 & 0.1342 \\ 
  $\gamma$: Seniors (65+) & 0.0015 & -0.0568 & 0.0598 &   & 0.0545 & -0.0145 & 0.1235 \\ 
  $\gamma$: Black population share & 0.0099 & -0.0051 & 0.0248 &   & 0.0011 & -0.0149 & 0.0172 \\ 
  $\gamma$: Asian population share & -0.0644 & -0.0821 & -0.0467 &   & -0.0397 & -0.0582 & -0.0212 \\ 
  $\gamma$: Hispanic population share & -0.0074 & -0.0229 & 0.0081 &   & -0.0015 & -0.0185 & 0.0156 \\ 
  $\gamma$: Other Minorities & 0.1629 & -0.0725 & 0.3983 &   & 0.0723 & -0.1840 & 0.3286 \\ 
  Accessibility & 0.6575 & 0.5348 & 0.7801 &   & 0.2306 & 0.0907 & 0.3705 \\ 
  $\lambda$: spatial correlation &  &  &  &   & 0.6854 & 0.6459 & 0.7249 \\

@@

Increased physical activity and decreased obesity rates are critical
measures of improvement in public health. Although many have theorized
the link between park access and public health, previous literature has
produced mixed findings, perhaps owing to the wide range of
accessibility definitions. Using New York City as a case study, we
presented a holistic and flexible measurement for park accessibility
that compares the continuous distance to all parks in the region,
weighted against the size of the park and its other amenities, with
weights determinable through revealed preferences. In terms of physical
activity, we found a positive relationship where the least
park-accessible tracts have an expected physical activity participation
rate between 0.06 and 1.8 percentage points lower than the most
accessible tracts. We also found a suggestive though not significant
relationship between park accessibility and expected obesity rates
\emph{in addition to} the effect of physical activity participation.

>>>>

\begin{table*}
\caption{Comparison of SLX and SDEM Coefficients}
\label{tab:slx-sdem}
\begin{tabular}{l d{2.4} d{2.4} d{2.4} c d{2.4} d{2.4} d{2.4} }
\toprule
 & \multicolumn{3}{c}{SLX} & & \multicolumn{3}{c}{SDEM} \\
\cmidrule{2-4}\cmidrule{6-8}
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.} & 
 & \multicolumn{1}{c}{Estimate} & \multicolumn{2}{c}{95\% Conf.Int.}\\
\midrule
% latex table generated in R 4.0.0 by xtable 1.8-4 package
% Fri Sep 25 10:52:15 2020
 (Intercept) & -15.8791 & -26.8066 & -4.9515 &   & -24.7352 & -41.5391 & -7.9312 \\ 
  log(Density) & 0.0502 & -0.1280 & 0.2284 &   & 0.1760 & 0.0371 & 0.3150 \\ 
  log(Income) & 5.8667 & 5.3476 & 6.3859 &   & 6.0134 & 5.5901 & 6.4368 \\ 
  Fulltime & 0.1157 & 0.0944 & 0.1369 &   & 0.1271 & 0.1086 & 0.1455 \\ 
  College-educated & 0.0029 & -0.0253 & 0.0311 &   & 0.0077 & -0.0156 & 0.0310 \\ 
  Single Adults & -0.0435 & -0.0623 & -0.0247 &   & -0.0365 & -0.0528 & -0.0201 \\ 
  Youth (0-17) & -0.1528 & -0.1844 & -0.1211 &   & -0.1320 & -0.1585 & -0.1055 \\ 
  Young adults (18-34) & 0.0245 & -0.0001 & 0.0492 &   & 0.0292 & 0.0092 & 0.0493 \\ 
  Seniors (65+) & 0.0108 & -0.0212 & 0.0427 &   & 0.0386 & 0.0112 & 0.0659 \\ 
  Black population share & -0.0471 & -0.0589 & -0.0353 &   & -0.0499 & -0.0586 & -0.0412 \\ 
  Asian population share & -0.0684 & -0.0828 & -0.0540 &   & -0.0764 & -0.0870 & -0.0659 \\ 
  Hispanic population share & -0.1003 & -0.1126 & -0.0881 &   & -0.1023 & -0.1114 & -0.0932 \\ 
  Other Minorities & 0.0341 & -0.0661 & 0.1343 &   & 0.0047 & -0.0837 & 0.0931 \\ 
  $\gamma$: log(Density) & 1.2513 & 0.9367 & 1.5659 &   & 1.0873 & 0.7062 & 1.4684 \\ 
  $\gamma$: log(Income) & 2.3641 & 1.5266 & 3.2016 &   & 1.9780 & 0.9563 & 2.9997 \\ 
  $\gamma$: Fulltime & -0.1346 & -0.1749 & -0.0943 &   & -0.0411 & -0.0882 & 0.0059 \\ 
  $\gamma$: College-educated & -0.1201 & -0.1692 & -0.0711 &   & -0.0704 & -0.1296 & -0.0112 \\ 
  $\gamma$: Single Adults & -0.0520 & -0.0857 & -0.0183 &   & -0.0094 & -0.0514 & 0.0325 \\ 
  $\gamma$: Youth (0-17) & -0.1013 & -0.1594 & -0.0432 &   & -0.0128 & -0.0815 & 0.0559 \\ 
  $\gamma$: Young adults (18-34) & 0.1131 & 0.0706 & 0.1555 &   & 0.0838 & 0.0334 & 0.1342 \\ 
  $\gamma$: Seniors (65+) & 0.0015 & -0.0568 & 0.0598 &   & 0.0545 & -0.0145 & 0.1235 \\ 
  $\gamma$: Black population share & 0.0099 & -0.0051 & 0.0248 &   & 0.0011 & -0.0149 & 0.0172 \\ 
  $\gamma$: Asian population share & -0.0644 & -0.0821 & -0.0467 &   & -0.0397 & -0.0582 & -0.0212 \\ 
  $\gamma$: Hispanic population share & -0.0074 & -0.0229 & 0.0081 &   & -0.0015 & -0.0185 & 0.0156 \\ 
  $\gamma$: Other Minorities & 0.1629 & -0.0725 & 0.3983 &   & 0.0723 & -0.1840 & 0.3286 \\ 
  Accessibility & 0.6575 & 0.5348 & 0.7801 &   & 0.2306 & 0.0907 & 0.3705 \\ 
  $\lambda$: spatial correlation &  &  &  &   & 0.6854 & 0.6459 & 0.7249 \\

@@

Increased physical activity and decreased obesity rates are critical
measures of improvement in public health. Although many have theorized
the link between park access and public health, previous literature has
produced mixed findings, perhaps owing to the wide range of
accessibility definitions. Using New York City as a case study, we
presented a holistic and flexible measurement for park accessibility
that compares the continuous distance to all parks in the region,
weighted against the size of the park and its other amenities, with
weights determinable through revealed preferences. In terms of physical
activity, we found a positive relationship where the least
park-accessible tracts have an expected physical activity participation
rate between \(0.3\) and \(2.1\) percentage points lower than the most
accessible tracts. We also found a suggestive though not significant
relationship between park accessibility and expected obesity rates
\emph{in addition to} the effect of physical activity participation.

### 2020-09-25 manuscript/submission/manuscript.tex
Add sentence about putting amenities into utility in contribution
Also, minor typo correction

\textcolor{blue}{To our knowledge, there has been no previous attempt to examine access
to parks in a metropolitan area using a utility-based definition as the
primary indicator of access. Consequently, there has also been no
previous effort to examine the effect of park access on health outcomes
using this theoretically attractive technique. In this study, we first
develop a utility-based measure of park access in New York City, and
then apply this measure to study the effect of access to public parks on
aggregate health outcomes.}

@@

Increased physical activity and decreased obesity rates are critical
measures of improvement in public health. Although many have theorized
the link between park access and public health, previous literature has
produced mixed findings, perhaps owing to the wide range of
accessibility definitions. Using New York City as a case study, we
presented a holistic and flexible measurement for park accessibility
that compares the continuous distance to all parks in the region,
weighted against the size of the park and its other amenities, with
weights determinable through revealed preferences. In terms of physical
activity, we found a positive relationship where the least
park-accessible tracts have an expected physical activity participation
rate between 0.06 and 1.8 percentage points lower than the most
accessible tracts. We also found a suggestive though not significant
relationship between park accessibility and expected obesity rates
\emph{in addition to} the effect of physical activity participation.

>>>>

\textcolor{blue}{To our knowledge, there has been no previous attempt to examine access
to parks in a metropolitan area using a utility-based definition as the
primary indicator of access. Consequently, there has also been no
previous effort to examine the effect of park access on health outcomes
using this theoretically attractive technique. In this study, we first
develop a utility-based measure of park access in New York City, and
then apply this measure to study the effect of access to public parks on
aggregate health outcomes. In this application, we demonstrate how additional
park amenities including park size, sport facilities, playgrounds, and trails 
can be incorporated in a more complete and informative measure of access beyond simple proximity.}

@@

Increased physical activity and decreased obesity rates are critical
measures of improvement in public health. Although many have theorized
the link between park access and public health, previous literature has
produced mixed findings, perhaps owing to the wide range of
accessibility definitions. Using New York City as a case study, we
presented a holistic and flexible measurement for park accessibility
that compares the continuous distance to all parks in the region,
weighted against the size of the park and its other amenities, with
weights determinable through revealed preferences. In terms of physical
activity, we found a positive relationship where the least
park-accessible tracts have an expected physical activity participation
rate between $0.3$ and $2.1$ percentage points lower than the most
accessible tracts. We also found a suggestive though not significant
relationship between park accessibility and expected obesity rates
\emph{in addition to} the effect of physical activity participation.
